\documentclass[12pt]{book}
\usepackage{amsmath, amssymb}
\usepackage{hyperref}

\begin{document}

\frontmatter
\title{Glossary of Linear Algebra --- MIT 18.06}
\author{}
\date{}
\maketitle

\tableofcontents

\mainmatter

\chapter{Adjacency Matrix of a Graph}
Let $G = (V,E)$ be a graph with vertices
\[
V = \{v_1, v_2, \dots, v_n\}.
\]

The \emph{adjacency matrix} of $G$ is the $n \times n$ matrix
\[
A = (a_{ij}),
\]
defined by
\[
a_{ij} =
\begin{cases}
1, & \text{if there is an edge between } v_i \text{ and } v_j, \\
0, & \text{otherwise}.
\end{cases}
\]

\begin{itemize}
  \item If $G$ is undirected, then $A$ is symmetric ($A = A^T$).
  \item If $G$ is directed, $A$ need not be symmetric: $a_{ij}=1$ if there is an arrow from $v_i$ to $v_j$.
  \item For weighted graphs, $a_{ij}$ equals the edge weight.
\end{itemize}



\chapter{Affine Transformation}
$T(v) = Av + v_0$ = linear transformation plus shift.

\chapter{Associative Law}
$(AB)C = A(BC)$. Parentheses can be removed to leave $ABC$.

\chapter{Augmented Matrix}
$[A \mid b]$. $Ax = b$ is solvable when $b$ is in the column space of $A$; then $[A \mid b]$ has the same rank as $A$. Elimination on $[A \mid b]$ keeps equations correct.

\chapter{Back Substitution}
Upper triangular systems are solved in reverse order $x_n$ to $x_1$.

\chapter{Basis}
Independent vectors $v_1,\dots,v_d$ whose linear combinations give every $v$ in $V$. A vector space has many bases.

\chapter{Big Formula for Determinants}
$\det(A)$ is a sum of $n!$ terms, one for each permutation $P$ of the columns. That term is the product along the diagonal of the reordered matrix, times $\det(P) = \pm 1$.

\chapter{Block Matrix}
A matrix partitioned into blocks by cuts between rows and/or columns. Block multiplication $AB$ is allowed if block shapes permit (columns of $A$ and rows of $B$ in matching blocks).

\chapter{Cayley--Hamilton Theorem}
$p(\lambda) = \det(A - \lambda I)$ has $p(A) = 0$.

\chapter{Change of Basis Matrix}
Old basis vectors $v_j$ are linear combinations of new basis vectors $w_i$. Coordinates transform as $d = M c$.

\chapter{Characteristic Equation}
$\det(A - \lambda I) = 0$. Roots $\lambda$ are the eigenvalues of $A$.

\chapter{Cholesky Factorization}
$A = C C^T = (L \sqrt{D})(L \sqrt{D})^T$ for positive definite $A$.

\chapter{Circulant Matrix}
Constant diagonals wrap around like a cyclic shift $S$. Every circulant is $c_0 I + c_1 S + \cdots + c_{n-1} S^{n-1}$.

\chapter{Cofactor}
Remove row $i$, column $j$; multiply determinant by $(-1)^{i+j}$.

\chapter{Column Picture of $Ax = b$}
$b$ is a combination of the columns of $A$. Solvable only when $b \in \mathcal{C}(A)$.

\chapter{Column Space}
$\mathcal{C}(A)$ = all linear combinations of the columns of $A$.

\chapter{Commuting Matrices}
$AB = BA$. If diagonalizable, they share $n$ eigenvectors.

\chapter{Companion Matrix}
Put $c_1,\dots,c_n$ in last row, ones on subdiagonal. Its characteristic polynomial is $c_1 + c_2 \lambda + \cdots$.

\chapter{Complete Solution}
$x = x_p + x_n$. Particular solution $x_p$ plus nullspace solution $x_n$.

\chapter{Complex Conjugate}
For $z=a+ib$, $\bar z = a - ib$. $z \bar z = |z|^2$.

\chapter{Condition Number}
$\kappa(A) = \|A\|\|A^{-1}\| = \sigma_{\max}/\sigma_{\min}$. Measures sensitivity of $Ax=b$ to changes in $b$.

\chapter{Conjugate Gradient Method}
Iterative method for positive definite $Ax=b$ minimizing $\tfrac{1}{2}x^T A x - x^T b$ over Krylov subspaces.

\chapter{Covariance Matrix}
$\Sigma_{ij} = \mathbb{E}[x_i x_j]$ for centered random variables. $\Sigma$ is positive semidefinite.

\chapter{Cramer’s Rule}
For invertible $A$, $x_j = \det(B_j)/\det(A)$ where $B_j$ replaces column $j$ of $A$ with $b$.

\chapter{Cross Product}
In $\mathbb{R}^3$, $u \times v$ is perpendicular to $u,v$, with length $\|u\|\|v\|\sin\theta$.

\chapter{Cyclic Shift}
Permutation matrix shifting entries. Eigenvalues are $n$th roots of unity; eigenvectors are columns of Fourier matrix.

\chapter{Determinant}
$\det(I)=1$, row swap changes sign, linear in rows. $\det(AB)=\det A \det B$, $\det A^T = \det A$.

\chapter{Diagonal Matrix}
$d_{ij}=0$ if $i\neq j$.

\chapter{Diagonalizable Matrix}
Has $n$ independent eigenvectors $\Rightarrow S^{-1} A S = \Lambda$ diagonal.

\chapter{Diagonalization}
$A=S\Lambda S^{-1}$, with $\Lambda$ diagonal of eigenvalues.

\chapter{Dimension}
$\dim V$ = number of vectors in a basis.

\chapter{Distributive Law}
$A(B+C)=AB+AC$.

\chapter{Dot Product}
$x^T y = \sum x_i y_i$. Zero dot product $\Leftrightarrow$ orthogonal.

\chapter{Echelon Matrix}
Row-echelon form: pivots staggered, zero rows at bottom.

\chapter{Eigenvalue and Eigenvector}
$Ax=\lambda x$, $x\neq 0$, $\det(A-\lambda I)=0$.

\chapter{Elimination}
Row operations reducing $A$ to triangular form $U$ (or reduced form $R$). $A=LU$.

\chapter{Ellipse}
$x^T A x=1$ for positive definite $A$. Eigenvectors give ellipse axes.

\chapter{Exponential of a Matrix}
$e^{At} = I + At + \frac{(At)^2}{2!}+\cdots$. Derivative: $A e^{At}$. Solves $u'(t)=Au$.

\chapter{Four Fundamental Subspaces}
For $A \in \mathbb{R}^{m \times n}$:
\[
\mathcal{C}(A), \quad \mathcal{N}(A), \quad \mathcal{C}(A^T), \quad \mathcal{N}(A^T).
\]
These subspaces are mutually orthogonal in pairs and fundamental to linear algebra.

\chapter{Fourier Series}
Representation of periodic functions $f(x)$ as sums of sines and cosines:
\[
f(x) \sim a_0 + \sum_{n=1}^\infty (a_n \cos nx + b_n \sin nx).
\]

\chapter{Fourier Transform}
Integral transform converting functions between time and frequency domain:
\[
\hat f(\xi) = \int_{-\infty}^\infty f(x) e^{-2\pi i x \xi} dx.
\]

\chapter{Frobenius Norm}
\[
\|A\|_F = \left( \sum_{i,j} |a_{ij}|^2 \right)^{1/2}.
\]

\chapter{Fundamental Theorem of Linear Algebra}
The four subspaces are related by orthogonal complements:
\[
\mathcal{C}(A)^\perp = \mathcal{N}(A^T), \quad \mathcal{C}(A^T)^\perp = \mathcal{N}(A).
\]

\chapter{Gauss--Jordan Elimination}
Row reduction producing reduced row echelon form (RREF).

\chapter{Gaussian Elimination}
Row operations to reduce $A$ to upper triangular form $U$.

\chapter{General Linear Group}
$GL(n,\mathbb{R})$ = set of invertible $n \times n$ real matrices.

\chapter{General Solution}
Complete solution of $Ax=b$: $x_p + x_n$.

\chapter{Geometric Multiplicity}
Dimension of eigenspace corresponding to an eigenvalue.

\chapter{Gram--Schmidt Process}
Algorithm for orthogonalizing a set of vectors.

\chapter{Gram Matrix}
$G = A^T A$. Positive semidefinite.

\chapter{Graph Laplacian}
$L = D - A$ where $D$ is degree matrix, $A$ adjacency matrix.

\chapter{Greatest Integer Function}
$\lfloor x \rfloor$ = largest integer $\leq x$.

\chapter{Hadamard Matrix}
Square $\pm 1$ matrix with orthogonal rows.

\chapter{Hadamard Product}
Entrywise product $(A \circ B)_{ij} = a_{ij} b_{ij}$.

\chapter{Hermitian Matrix}
$A = A^*$. Eigenvalues are real.

\chapter{Hilbert Matrix}
$H_{ij} = \tfrac{1}{i+j-1}$. Ill-conditioned.

\chapter{Householder Matrix}
$H = I - 2uu^T$ for unit vector $u$. Reflects across hyperplane.

\chapter{Identity Matrix}
$I_n$ = $n \times n$ matrix with $1$ on diagonal, $0$ elsewhere.

\chapter{Image}
$\mathrm{Im}(A) = \mathcal{C}(A)$.

\chapter{Inner Product}
$\langle x,y \rangle = x^Ty$ (or $y^* x$ for complex).

\chapter{Inverse Matrix}
$A^{-1}$ satisfies $AA^{-1}=I$.

\chapter{Inverse of a Partitioned Matrix}
Blockwise formulas using Schur complement.

\chapter{Isomorphism}
Linear bijection between vector spaces.

\chapter{Iteration Matrix}
$M$ used in iterative methods: $x^{(k+1)} = M x^{(k)} + c$.

\chapter{Jordan Canonical Form}
$A = P J P^{-1}$, $J$ block diagonal with Jordan blocks.

\chapter{Kernel}
Nullspace $\mathcal{N}(A)$.

\chapter{Krylov Subspace}
$\mathcal{K}_k(A,r) = \mathrm{span}\{r, Ar, A^2 r, \dots, A^{k-1}r\}$.

\chapter{Least Squares Problem}
Solve $\min \|Ax-b\|_2$. Normal equations: $A^TAx=A^Tb$.

\chapter{LU Factorization}
$A=LU$ with lower $L$, upper $U$.

\chapter{Markov Matrix}
Square matrix with nonnegative entries and each column summing to 1.

\chapter{Matrix Norm}
Any function $\|\cdot\|$ satisfying norm properties extended to matrices.

\chapter{Minimal Polynomial}
Smallest degree monic polynomial $p$ such that $p(A)=0$.

\chapter{Moore--Penrose Pseudoinverse}
$A^+$ gives least squares solution: $x=A^+ b$.

\chapter{Normal Equations}
For least squares: $A^T A x = A^T b$.

\chapter{Normal Matrix}
$A$ satisfies $AA^* = A^* A$. Diagonalizable by a unitary matrix.

\chapter{Norm of a Vector}
$\|x\|_p = (|x_1|^p + \cdots + |x_n|^p)^{1/p}$ for $p \geq 1$.

\chapter{Nullspace}
$\mathcal{N}(A) = \{x : Ax=0\}$.

\chapter{Nullity}
Dimension of $\mathcal{N}(A)$.

\chapter{Numerical Rank}
Effective rank determined by singular values above tolerance.

\chapter{Orthogonal Complement}
$V^\perp = \{x : x^Tv=0 \ \forall v\in V\}$.

\chapter{Orthogonal Matrix}
$Q^T Q = I$. Inverse equals transpose.

\chapter{Orthogonal Projection}
$P = QQ^T$ for orthonormal $Q$.

\chapter{Orthogonal Subspaces}
$U \perp V$ if $u^Tv=0$ for all $u \in U$, $v \in V$.

\chapter{Orthogonal Vectors}
$x^Ty=0$.

\chapter{Orthonormal Vectors}
$x^Ty = 0$ if $x \neq y$, and $\|x\|=1$.

\chapter{Outer Product}
$uv^T$.

\chapter{Partial Pivoting}
Row swaps in elimination to control growth.

\chapter{Permutation Matrix}
Matrix obtained from $I$ by permuting rows.

\chapter{Pivot}
First nonzero entry in each row after elimination.

\chapter{Poisson Matrix}
Tridiagonal with $2$ on diagonal and $-1$ on off-diagonal, from discretized Laplace operator.

\chapter{Polar Decomposition}
$A = Q H$ with $Q$ orthogonal, $H$ symmetric positive semidefinite.

\chapter{Positive Definite Matrix}
$x^T A x > 0$ for all $x \neq 0$.

\chapter{Power Method}
Iterative method converging to dominant eigenvector.

\chapter{Principal Axis Theorem}
Real symmetric $A$ diagonalizable by orthogonal $Q$: $A=Q\Lambda Q^T$.

\chapter{Projection Matrix}
$P^2=P$, $\mathcal{C}(P)$ is subspace onto which vectors are projected.

\chapter{Pseudoinverse}
See Moore–Penrose pseudoinverse.

\chapter{QR Factorization}
$A=QR$ with $Q$ orthogonal, $R$ upper triangular.

\chapter{Quadratic Form}
$x^T A x$.

\chapter{Range}
Column space.

\chapter{Rank}
Dimension of $\mathcal{C}(A)$.

\chapter{Rank--Nullity Theorem}
$\dim\mathcal{C}(A) + \dim\mathcal{N}(A) = n$.

\chapter{Rayleigh Quotient}
$R(x) = \dfrac{x^T A x}{x^T x}$.

\chapter{Reduced Row Echelon Form}
Unique row echelon form with pivot 1s and zeros above and below.

\chapter{Reflection Matrix}
$H=I-2uu^T$. Orthogonal, symmetric.

\chapter{Ritz Values}
Eigenvalues of projected matrix $Q^T A Q$.

\chapter{Row Picture of $Ax=b$}
Each equation is a hyperplane; solution is intersection.

\chapter{Row Space}
$\mathcal{C}(A^T)$.

\chapter{Schur Complement}
If $A=\begin{bmatrix} B & C \\ D & E\end{bmatrix}$ with $B$ invertible, then $S=E-DB^{-1}C$.

\chapter{Schur’s Theorem}
Every square $A$ is unitarily similar to upper triangular.

\chapter{Similarity}
$A$ and $B$ are similar if $A=SBS^{-1}$.

\chapter{Singular Matrix}
$\det A=0$.

\chapter{Singular Value Decomposition (SVD)}
$A=U\Sigma V^T$ with orthogonal $U,V$.

\chapter{Singular Values}
Square roots of eigenvalues of $A^TA$.

\chapter{Skew-Symmetric Matrix}
$A^T=-A$. Eigenvalues are purely imaginary.

\chapter{Spectral Decomposition}
$A=Q\Lambda Q^T$ for symmetric $A$.

\chapter{Spectral Radius}
$\rho(A)=\max |\lambda|$ over eigenvalues.

\chapter{Spectrum}
Set of eigenvalues.

\chapter{Square Matrix}
$n \times n$ matrix.

\chapter{Stability}
Bounded perturbations lead to bounded solutions.

\chapter{Stochastic Matrix}
Square with nonnegative entries, columns sum to 1.

\chapter{Subspace}
Subset closed under addition and scalar multiplication.

\chapter{Symmetric Matrix}
$A=A^T$. Real eigenvalues, orthogonal eigenvectors.

\chapter{Toeplitz Matrix}
Constant diagonals.

\chapter{Trace}
$\mathrm{tr}(A)=\sum a_{ii}$. Invariant under similarity.

\chapter{Triangular Matrix}
Upper or lower triangular.

\chapter{Tridiagonal Matrix}
Nonzeros only on diagonal and first off-diagonals.

\chapter{Unitary Matrix}
$U^*U=I$.

\chapter{Vandermonde Matrix}
Rows are powers of given numbers.

\chapter{Vector Space}
Set closed under vector addition and scalar multiplication.

\backmatter

\end{document}
